---
title: "Case study: nimotuzumab target-mediated disposition"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Case study: nimotuzumab target-mediated disposition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
extdata <- function(f) read.csv(system.file("extdata", f, package = "pkident"), check.names = FALSE)
```

```{r setup}
library(pkident)
```

**Question.** Which parameters of a full target-mediated drug disposition
(TMDD) model are determined by the public `nimoData` observations
(nlmixr2data; source: Rodriguez-Vera et al. 2015), and which added
measurement would determine the rest?

The model fit and the likelihood profiles take about half an hour, so this
vignette shows their code without running it and reads the stored results
(`inst/extdata`), produced by `case-studies/nimotuzumab-tmdd.R` in the source
repository. The structural analysis runs here.

## Data and model

`nimoData` has 12 patients at 50–400 mg, each with ten infusions and 321
free-drug observations in total. DV is the log concentration; its unit is not
documented, so the volume is in dose units per concentration unit. The full
TMDD model has drug amount `A`, free target `R` (at its baseline `r0` before
the first dose) and complex `P`:

```{r}
m <- pkpd_model(
  "d/dt(A) = -cl/v*A - v*(kon*A/v*R - koff*P)
   d/dt(R) = r0*kdeg - kdeg*R - kon*A/v*R + koff*P
   d/dt(P) = kon*A/v*R - koff*P - kint*P
   cp = A/v
   rtot = R + P",
  outputs = c("cp", "rtot", "R", "P"),
  doses = bolus_dose("A", "DOSE"), initial = list(R = "r0"), known = "DOSE"
)
```

## 1. Structure is not the obstacle

With free drug observed alone, all seven parameters are locally identifiable:

```{r}
m_free <- m
m_free$outputs <- m$outputs["cp"]
structural_identifiability(m_free)
```

## 2. What these data determine

The model was fitted with nlmixr2 (FOCEi; between-subject variability on
`cl` and `v`; additive error on the log scale):

```{r, eval = FALSE}
fit <- nlmixr2est::nlmixr2(tmdd, nimoData, est = "focei")
prof <- classify_profiles(fit, structural = structural_identifiability(m_free),
                          map = c(lcl = "cl", lv = "v", lkon = "kon", lkoff = "koff",
                                  lkdeg = "kdeg", lr0 = "r0", lkint = "kint"))
```

```{r}
fit <- extdata("nimotuzumab-fit.csv")
knitr::kable(fit[, c("parameter", "Estimate", "SE", "Back-transformed")], digits = 4)
prof <- extdata("nimotuzumab-profiles.csv")
knitr::kable(prof[, c("parameter", "estimate", "status", "shape", "lower", "upper", "interpretation")],
             digits = 2)
```

Clearance, volume and `koff` are bounded. `kon`, `kdeg`, `r0` and `kint` are
flat over a factor of 7.4 in each direction: structurally identifiable, but not
determined by these data. The standard errors from nlmixr2's covariance step
(0.17–0.48 on the log scale for these four) do not describe this.

```{r, fig.width = 7, fig.height = 4}
pts <- extdata("nimotuzumab-profile-points.csv")
op <- par(mfrow = c(2, 4), mar = c(4, 4, 2, 1))
for (p in unique(pts$parameter)) {
  d <- pts[pts$parameter == p, ]
  plot(d$value, d$dofv_from_minimum, type = "b", pch = 19, xlab = p, ylab = "dOFV",
       ylim = c(0, max(8, d$dofv_from_minimum)), main = p)
  abline(h = 3.84, lty = 2)
}
par(op)
```

## 3. Which measurement would determine the rest

Expected RSEs at the fitted values for the actual design (each patient's own
infusions and sampling times) and for candidates with an added measurement at
the same times, or with an added 10 mg group. Target assays are assumed to
have an error of 0.2 on the log scale.

```{r, eval = FALSE}
cmp <- compare_designs(
  m,
  candidates = list(current = current, `+ total target` = with_output("rtot"),
                    `+ free target` = with_output("R"), `+ complex` = with_output("P"),
                    `+ 10 mg group` = c(current, low_dose)),
  values = values,
  error = list(cp = residual_error(exp = 0.668), rtot = residual_error(exp = 0.2),
               R = residual_error(exp = 0.2), P = residual_error(exp = 0.2))
)
```

```{r}
p <- extdata("nimotuzumab-designs-parameters.csv")
tab <- reshape(p[, c("candidate", "parameter", "rse_percent")], idvar = "parameter",
               timevar = "candidate", direction = "wide")
names(tab) <- sub("rse_percent.", "", names(tab), fixed = TRUE)
knitr::kable(tab, digits = 1, row.names = FALSE, caption = "Expected RSE (%)")
s <- extdata("nimotuzumab-designs-summary.csv")
knitr::kable(s[, c("candidate", "outputs", "observations", "n_identifiable", "n_parameters")])
```

**Answer.** Measuring total target at the same times would determine `kon`,
`kdeg` and `r0` as well (six of seven parameters with RSE ≤ 30%); free target
determines five; a lower dose group does not help. No candidate determines
`kint`: at the estimate, internalisation of the complex barely affects any
output.

The expected RSEs (fixed effects of a typical patient, local approximation)
and the profiles (the actual mixed-effects fit) agree on which parameters are
determined; they differ in degree, for example for `koff`.
