---
title: "Structural identifiability of reference PK/PD models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Structural identifiability of reference PK/PD models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(pkident)
```

pkident decides **local** structural identifiability: whether the unknown
parameters of an ODE model are determined, in a neighbourhood of their values,
by continuous noise-free observation of the outputs after the known doses. Two
independent methods are used and must agree:

* Lie derivatives of the outputs evaluated at the known initial state (the
  Taylor-series approach), which keeps the information that a dose gives
  about volumes;
* the rank of the output sensitivity matrix, solved with rxode2.

## A model, step by step

A one-compartment model with first-order absorption and unknown
bioavailability `F`:

```{r}
m <- pkpd_model(
  "d/dt(depot)   = -ka*depot
   d/dt(central) = ka*depot - CL/V*central
   cp            = central/V",
  outputs = "cp",
  doses   = bolus_dose("depot", "F*DOSE"),
  known   = "DOSE"
)
m
```

```{r}
r <- structural_identifiability(m)
r
```

`CL`, `F` and `V` are not identifiable alone; the reported combinations span
the identifiable ones. The familiar `V/F` is among them:

```{r}
as_table(r, "combinations")
as_table(r, "agreement")
```

## The acceptance set

The package ships ten reference models whose expected results were fixed from
primary literature (Janzén et al. 2016, Front Physiol 7:590) or derived
analytically before any code was written, plus a variant of A10 with total
receptors known.

```{r}
reference_models()
```

```{r}
decide <- function(id) {
  m <- reference_model(id)
  r <- structural_identifiability(m)
  p <- r$parameters
  data.frame(
    model = id,
    decision = r$decision,
    methods_agree = sum(r$agreement$agree),
    identifiable = paste(p$parameter[p$status == "identifiable"], collapse = ", "),
    not_alone = paste(p$parameter[p$status != "identifiable"], collapse = ", "),
    combinations = paste(r$combinations$combination, collapse = ", "),
    matches_literature = setequal(p$parameter[p$status == "identifiable"],
                                  attr(m, "expected")$identifiable)
  )
}
results <- do.call(rbind, lapply(reference_models()$id, decide))
knitr::kable(results)
```

Every decision matches the expected result, and the two methods agree at all
five random points for every model.

## Reading the results

* `identifiable`: locally identifiable. It may still have a finite number of
  alternative values (A3, the oral model with `F` known, has the flip-flop
  solution of absorption and elimination).
* `combination_only`: not identifiable alone, but identifiable within a
  reported combination.
* `non_identifiable`: not identifiable, and in no reported combination.
* `unresolved`: the computation could not decide (limits reached, poor
  conditioning, or the two methods disagree). pkident never turns an
  undecided case into a decision.

See `validation/reference-models.md` in the source repository for the
derivations and the full validation record.
